---
title: "linelist_aggregation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{linelist_aggregation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```




# Outline of the report

# Step 2: Line list aggregation

This report explains necessary data preparation processes after importing a line list data frame.\

**Input**

* Case data in a line list format (**`cases_clean.rds`**) produced in the previous step ([link](...)).


**output**

* Case incidence grouped by time interval (**`cases.rds`**)


## Loading libraries

The following code loads required packages; missing packages will be installed
automatically, but will require a working internet connection for the
installation to be successful.

```{r setup, message = FALSE}
library(dplyr) # For data wrangling
library(incidence2) # For generating incidence data
```


## Input data

In this example, we will use a line list data extracted from Go.Data API.
While the example data set is already cleaned, it is important to make sure you have applied necessary data cleaning steps such as:  

* Column names are cleaned and standardized
* Data types are properly assigned
* Missing / duplicated values are treated
* Categorical values are cleaned



```{r}
# Go.Data extract as an example
cases_clean <- readRDS("data/Go.Data exports/cases_clean.rds")

head(cases_clean)
```


## Format the date columns properly

You will need one date index column to produce the aggregated incidence output.
While this is optional, in this example, since we have `nrow(cases_clean[is.na(cases_clean$date_of_onset),])` out of `nrow(cases_clean)` cases without any `date_of_osnet` values, we will compliment this with the `date_of_reporting` value.

```{r}
cases_clean <- cases_clean %>%
  mutate(
    date_of_onset = case_when(
      is.na(date_of_onset) ~ date_of_reporting,
      TRUE ~ date_of_onset)
  )
```



## Apply necessary aggregation and rename

We will convert this into a time-indexed grouped data frame with three columns:  

* `date_index` - This is the time index for your analysis (eg. date of infection, onset of symptoms, reporting...)
* `group` - Any grouping variable you might need (eg. places, age groups)
* `count` - Counts of observations per time and group


There are two ways of achieving this step as follows.


### Option 1: Using `incidence()` function from the _incidence2_ package.
In this method, the output will be in an **incidence2** class object. With this, you can also simply plot your incidence data.
See more information on the _incidence2_ package [here](https://cran.r-project.org/web/packages/incidence2/vignettes/incidence2.html).

```{r}
# Generate the incidence object
cases <- incidence(cases_clean %>% filter(!is.na(admin_1_name)), 
                   groups = "admin_1_name", 
                   date_index = "date_of_onset")
  
# Rename necessary columns
cases <- cases %>%
  rename(group = admin_1_name)


head(cases)
```
```{r}
plot(cases)
```



### Option 2: Using `group_by()` function from the _dplyr_ package.

In this method, your output will be a data frame object.

```{r}
cases <- cases_clean %>% filter(!is.na(admin_1_name)) %>%
  group_by(date_index=date_of_onset, group=admin_1_name) %>%
  summarise(count = n())

head(cases)
```
