---
title: "Rt_EpiEstim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rt_EpiEstim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, setup, eval=FALSE}
library(EpiEstim)
library(ggplot2)
```

### Explanations

### Results

```{r, rt-wrappers-prep, eval=FALSE}

# Function to process output from `EpiEstim::estimate_R` and output a tibble
# with dates, mean R, and 95% credible intervals
#'
#' @param x An `estimate_R` object
#' @param incid (Optional) The incidence2 object on which R values have been
#'   estimated.
wrap_res <- function(x, incid = NULL) {
  stopifnot(inherits(x, "estimate_R"))
  out <- tibble::tibble(x$R)
  out <- dplyr::select(out,
    start = t_start,
    end = t_end,
    mean = `Mean(R)`,
    sd = `Std(R)`,
    median = `Median(R)`,
    lower = `Quantile.0.025(R)`,
    upper = `Quantile.0.975(R)`
  )
  if (!is.null(incid)) {
    stopifnot(inherits(incid, "incidence2"))
    dates <- unique(incidence2::get_dates(incid))
    out$start <- dates[out$start]
    out$end <- dates[out$end]
  }

  class(out) <- c("R_estimate", class(out))
  out
}
```


#### Serial interval settings

Here we setup the serial interval settings for EpiEstim to use, we will assume a
mean of 4.2 and standard deviation of 4.9, using the `"parametric_si"` method in
EpiEstim.
```{r, rt-estim-si, eval=FALSE}
# Run EpiEstim
config_epiestim <- EpiEstim::make_config(mean_si = 4.2,
                                         std_si = 4.9)
```

#### Global transmissibility

These analyses present results for the global incidence, i.e. without
stratification. Results include:
 
We now use EpiEstim to estimate Rt for the global incidence, i.e.
without stratification. As a default these will be estimated on weekly sliding
windows.

```{r, rt-estim-global, eval=FALSE}
res_epiestim_global <- cases_clean |>
  incidence2::regroup() |>
  incidence2::get_count_value() |>
  EpiEstim::estimate_R(method = "parametric_si",
                       config = config_epiestim) |>
  wrap_res(cases_clean)
```
We plot Rt, showing the median as a dark green line and the 95% credible
interval is represented by a pale green area. Values are plotted at the end date
for each of the weekly sliding windows.

```{r, rt-plot-global, eval=FALSE}
# Graph of all values over time
ggplot(res_epiestim_global, aes(x = end)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = pale_green) +
  geom_line(aes(y = median), color = dark_green) +
  geom_hline(yintercept = 1, color = dark_pink) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    x = "",
    y = "Instantaneous reproduction number (Rt)",
    title = "Estimates of Rt (EpiEstim)"
  )
```
We output a table giving the mean, median and 95% credible intervals, reporting
by the end date of each weekly sliding window.
```{r, rt-table-global, eval=FALSE}
# Table
res_epiestim_global |>
  tail(params$r_estim_window) |>
  dplyr::mutate(
    date = end,
    mean = round(mean, 2),
    median = round(median, 2),
    `95% ci` = sprintf(
      "[%1.2f ; %1.2f]",
      lower,
      upper
    )
  ) |>
  dplyr::select(date, mean, median, `95% ci`) |>
  dplyr::rename(
    "mean $R$" = mean,
    "median $R$" = median
  ) |>
  purrr::set_names(toupper) |>
  kableExtra::kbl() |>
  kableExtra::kable_paper("striped", font_size = 18, full_width = FALSE)
```

#### Transmissibility by group
We now run EpiEstim to estimate Rt again, but this time keep the data stratified
into groups and calculate Rt for each group based on the incidence in that
group.
```{r, rt-estim-group, eval=FALSE}
res_epiestim_group <- cases_clean |>
  tidyr::nest(data = c(incidence2::get_date_index_name(.),
                       incidence2::get_count_value_name(.))) |>
  dplyr::mutate(
    res_epiestim = map(data, ~ wrap_res(
      EpiEstim::estimate_R(.x$count,
                           method = "parametric_si",
                           config = config_epiestim),
      cases_clean)
    )
  ) |>
  tidyr::unnest(res_epiestim) |>
  dplyr::select(-data)
```

We plot Rt for each group, showing the median as a dark green line and the 95%
credible interval is represented by a pale green area. Values are plotted at the
end date for each of the weekly sliding windows.

```{r, rt-plot-group, eval=FALSE}
# Graph of all values over time
ggplot(res_epiestim_group, aes(x = end)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = pale_green) +
  geom_line(aes(y = median), color = dark_green) +
  geom_hline(yintercept = 1, color = dark_pink) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  facet_wrap(~ .data[[group_var]], ncol = 1) +
  labs(
    x = "",
    y = "Instantaneous reproduction number (Rt)",
    title = "Estimates of Rt (EpiEstim)"
  )
```

Here we produce a whiskers plot of the latest estimate of Rt (i.e. for the final
weekly rolling window) for each region, with dots representing the median and
the whiskers the 95% credible intervals.
```{r, rt-plot-estimates-group, eval=FALSE}
# Plot of latest estimates
res_epiestim_group |>
  dplyr::filter(end == max(end)) |>
  ggplot(aes(y = .data[[group_var]]), fill = custom_grey) +
    geom_point(aes(x = median), color = dark_green) +
    geom_errorbar(aes(xmin = lower, xmax = upper), color = dark_green) +
    geom_vline(xintercept = 1, color = dark_pink) +
    labs(
      title = "Latest estimates of Rt",
      subtitle = sprintf(
        "As of %s",
        format(max(get_dates(cases_clean)), "%d %B %Y")
      ),
      y = "",
      x = "Instantaneous Reproduction Number (Rt)"
    )
```
Now we present the estimates of Rt by region for the final weekly rolling window
in a table, giving the mean, median and 95% credible interval.
```{r, rt-table-group, eval=FALSE}
res_epiestim_group |>
  dplyr::filter(end == max(end)) |>
  dplyr::mutate(
    mean = round(mean, 2),
    median = round(median, 2),
    `95% ci` = sprintf(
      "[%1.2f ; %1.2f]",
      lower,
      upper
    )
  ) |>
  dplyr::select(-c(sd, lower, upper, count_variable)) |>
  dplyr::rename(
    "mean $R$" = mean,
    "median $R$" = median
  ) %>%
  purrr::set_names(toupper) |>
  kableExtra::kbl() |>
  kableExtra::kable_paper("striped", font_size = 18, full_width = FALSE)
```
