---
title: "Rt_EpiEstim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rt_EpiEstim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(EpiEstim)
```

## _EpiEstim_ 

```{r lockfile-rt, include = FALSE, message = FALSE}
renv::use("EpiEstim@2.2-4")
```

```{r rt-load-pkg}
library(EpiEstim)
```

### Explanations

### Results

```{r, eval = params$use_epiparameter}
si_epiparameter <- epiparameter::epidist(
  params$epiparameter_pathogen,
  "serial_interval"
)
si_params <- si_epiparameter$param
si_dist <- si_epiparameter$dist
si_mean <- si_params[[1]]
si_sd <- si_params[[2]]
```

```{r, eval = !params$use_epiparameter}
si_mean <- params$si_mean
si_sd <- params$si_sd
si_dist <- params$si_dist
si_cv <- si_sd / si_mean
si_params <- epitrix::gamma_mucv2shapescale(mu = si_mean, cv = si_cv)
```

```{r}
# do.call() because we want to pass `si_params` names as the name of arguments.
# This is important if params in si_params are not the first arguments of the
# q...() function. For example, if a gamma distribution is define by its shape
# and scale, instead of shape and rate.
si <- do.call(
  distcrete::distcrete,
  c(
    list(si_dist,
      interval = 1,
      w = 0
    ),
    si_params
  )
)
si_x <- seq(1L, to = si$q(0.999), by = 1L)
```

```{r rt-wrappers-prep}
# Function to process a distrcrete object and output `si_discr` argument for
# EpiEstim::make_config

wrap_si <- function(x) {
  stopifnot(inherits(x, "distcrete"))
  max_x <- x$q(0.999)
  x <- si$d(seq_len(max_x))
  x[1] <- 0
  x <- x / sum(x)
  x
}


# Function to process output from `EpiEstim::estimate_R` and output a tibble
# with dates, mean R, and associated confidence intervals
#'
#' @param x An `estimate_R` object
#' @param incid (Optional) The incidence2 object on which R values have been
#'   estimated.
wrap_res <- function(x, incid = NULL) {
  stopifnot(inherits(x, "estimate_R"))
  out <- tibble::tibble(x$R)
  out <- dplyr::select(out,
    start = t_start,
    end = t_end,
    mean = `Mean(R)`,
    sd = `Std(R)`,
    median = `Median(R)`,
    lower = `Quantile.0.025(R)`,
    upper = `Quantile.0.975(R)`
  )
  if (!is.null(incid)) {
    stopifnot(inherits(incid, "incidence2"))
    dates <- unique(incidence2::get_dates(incid))
    out$start <- dates[out$start]
    out$end <- dates[out$end]
  }

  class(out) <- c("R_estimate", class(out))
  out
}

# Run EpiEstim
ee_config <- EpiEstim::make_config(si_distr = wrap_si(si))
```

#### Global transmissibility

These analyses present results for the global incidence, i.e. without
stratification. Results include:
 
* a graph of the estimates of $R_t$ of the observed data over time, with
associated 95% credibility intervals

* a table summarising these results for the last `r params$r_estim_window` days

```{r rt-estim-global}
res_epiestim_global <- dat_i_day %>%
  incidence2::regroup() %>%
  incidence2::get_count_value() %>%
  EpiEstim::estimate_R(config = ee_config) %>%
  wrap_res(dat_i_day)
```

```{r rt-plot-global}
# Graph of all values over time
ggplot(res_epiestim_global, aes(x = end)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = pale_green) +
  geom_line(aes(y = median), color = dark_green) +
  geom_hline(yintercept = 1, color = dark_pink) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
    x = "",
    y = "Instantaneous reproduction number (Rt)",
    title = "Estimates of Rt (EpiEstim)"
  )
```

```{r rt-table-global}
# Table
res_epiestim_global %>%
  tail(params$r_estim_window) %>%
  mutate(
    date = end,
    mean = round(mean, 2),
    median = round(median, 2),
    `95% ci` = sprintf(
      "[%1.2f ; %1.2f]",
      lower,
      upper
    )
  ) %>%
  dplyr::select(date, mean, median, `95% ci`) %>%
  rename(
    "mean $R$" = mean,
    "median $R$" = median
  ) %>%
  set_names(toupper) %>%
  kbl() %>%
  kable_paper("striped", font_size = 18, full_width = FALSE)
```

#### Transmissibility by group

```{r rt-estim-group}
# Get results by group, keeping only the last 14 days of data
res_epiestim_group <- dat_i_day %>%
  nest(data = c(incidence2::get_date_index_name(.),
                incidence2::get_count_value_name(.))) %>%
  mutate(
    res_epiestim = map(data, ~ wrap_res(
      EpiEstim::estimate_R(.x$count, config = ee_config),
      dat_i_day)
    )
  ) %>%
  unnest(res_epiestim) %>%
  dplyr::select(-data)
```

```{r rt-plot-group, fig.height = 5 / 3 * n_groups}
# Graph of all values over time
ggplot(res_epiestim_group, aes(x = end)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = pale_green) +
  geom_line(aes(y = median), color = dark_green) +
  geom_hline(yintercept = 1, color = dark_pink) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  facet_wrap(~ .data[[group_var]], ncol = 1) +
  labs(
    x = "",
    y = "Instantaneous reproduction number (Rt)",
    title = "Estimates of Rt (EpiEstim)"
  )
```

```{r rt-plot-estimates-group}
# Plot of latest estimates
res_epiestim_group %>%
  filter(end == max(end)) %>%
  ggplot(aes(y = .data[[group_var]]), fill = custom_grey) +
    geom_point(aes(x = median), color = dark_green) +
    geom_errorbar(aes(xmin = lower, xmax = upper), color = dark_green) +
    geom_vline(xintercept = 1, color = dark_pink) +
    labs(
      title = "Latest estimates of Rt",
      subtitle = sprintf(
        "As of %s",
        format(max(get_dates(dat_i_day)), "%d %B %Y")
      ),
      y = "",
      x = "Instantaneous Reproduction Number (Rt)"
    )
```

```{r rt-table-group}
res_epiestim_group %>%
  filter(end == max(end)) %>%
  mutate(
    mean = round(mean, 2),
    median = round(median, 2),
    `95% ci` = sprintf(
      "[%1.2f ; %1.2f]",
      lower,
      upper
    )
  ) %>%
  dplyr::select(-c(sd, lower, upper)) %>%
  rename(
    "mean $R$" = mean,
    "median $R$" = median
  ) %>%
  set_names(toupper) %>%
  kbl() %>%
  kable_paper("striped", font_size = 18, full_width = FALSE)
```
